\Large If the Null Fits, You Must Omit: Measurement Artifacts Undermine Group Polarization Replications

### 2025-12-05 GroupPolarizationStatmod/paper/method.tex
starting penultimate revision

\section{Method}\label{method}

A group polarization replication is proven undecidable, and therefore not
truly a replication, if we can find a model of \emph{mere agreement} that
appears to be polarized under that replication's experimental design. If we
cannot find such a model, we cannot determine a replication to be
undecidable. Mere agreement in this model is a dynamic where agreement
increases within the group, but extremism does not—in other words, A
\emph{null-polarization distributions} is a set of two Gaussian
distributions: one representing the pre-discussion opinions of the group, and
the other representing the post-discussion opinions. Our
\emph{null-polarization distributions} represent the process of the group
finding \emph{mere agreement}, which is when the average group opinion is
constant over time, but opinion variance decreases to represent agreement
(or, equivalently, \emph{consensus}~\cite{DeGroot1974}).

Null-polarization  We thus model opinions at each time, $t=\mathrm{pre}$ and
$t=\mathrm{post}$, as random variables drawn from a normal distribution with
mean $\mu_t$ and standard deviation $\sigma_t$.  we drop the time

The null-polarization distributions are represented by

in a simple, general way—i.e., a decrease in opinion variance but a static
opinion mean—we represent . In the general group polarization hypothesis, the
group is assumed to start out mostly agreeing in principle (i.e., mostly all
agree or disagree) but to different extremes. After discussion

Showing a replication is undecidable is sufficient to undermine it. However,
it does not tell us what is the probability that a particular experimental
design produces false positives, due to clipping or any other problem.

first develop a mathematical model of estimate the best-case scenario false
positive rate for each experimental design,

\subsection{Mathematical Model of Null-Polarization Measurement Artifacts}

Latent opinions are modeled as normally-distributed random variables.  
This is a simple, sufficiently general way to represent group-level changes
in opinions as encoded in the general hypothesis of group polarization, that
likeminded groups tend to become more extreme on average as they discuss.
For time, $t$, then, we represent each participant $i$'s opinion as 
\begin{equation} 
  \omega_{i,t} \sim \mathcal{N}(\mu_t, \sigma_t).
  \label{eq:opinionDistribution} 
\end{equation} 
\noindent

Group-level opinion change, then, can be represented as a change in opinion
mean, opinion variance, or both. Group polarization occurs 
when a group's average opinion increases over time. When likeminded people
interact, they tend to come to agree more over time—they \emph{find consensus}. 
By definition this means opinion variance decreases, 
whether the average changes or not.

We can define a null-polarization model, then, as one where the group's
average opinion stays constant, but the variance decreases: $\sigmapost
\leq \sigmapre$. Since the mean opinion is constant we just call it 
$\mu = \mupre = \mupost$. For conciseness, we define a null-polarization
model as a 3-tuple, $\nu = (\mu, \sigmapre, \sigmapost)$.

\begin{center}
\vspace{1.5em}
[ TABLE \ref{tab:latent-variables} ABOUT HERE ]
\vspace{1.5em}
\end{center}

\subsubsection{Simulated opinion measurements}

We simulate the collection of an \emph{ordinal distribution} of \emph{ordinal
observations}, as was performed in a prototypical group polarization experiment 
by integrating the latent opinion probability density function, $f(\omega)$,
over each bin's corresponding range of opinion values. For example, consider an
ordinal scale that runs from -3 (strongly disagree) to +3 (strongly agree)
(as used by \textcite{Moscovici1969}, for example).
the frequency with which the ``2'' bin is chosen is just the total
area underneath the probability density curve between 1.5 and 2.5.

Formally, the probability density function is
\begin{equation}
  f(\omega) = 
    \frac{1}{\sigma \sqrt{2 \pi}} e^{-\frac{(\omega - \mu)^2}{2 \sigma^2}}.
  \label{eq:normal_pdf}
\end{equation}
\noindent
To set up our measurement simulation, we will represent
oridnal bin values as $b_i$ with $i=1,\ldots,B$—i.e., there are $B$ bins. 
To create $B$ bins there must be $B+1$ \emph{thresholds} that define what
range of latent opinions correspond to which ordinal bin value.

\begin{center}
  \vspace{1.5em}
  [ TABLE \ref{tab:experiment-parameters} ABOUT HERE ]
  \vspace{1.5em}
\end{center}

All bins have a width of 1 in latent opinion space, except for the two most
extreme bins at opposite ends of the sentiment spectrum—this works because
bin values are always separated by 1. All continuous latent opinion values
$\omega < b_1 + \frac{1}{2} = \theta_1$ get mapped to the ordinal measurement
value $b_1$—so, $b_1$'s lower threshold is $\theta_0 = -\infty$.  Similarly,
all opinions $\omega > b_B - 0.5 = \theta_B$ are reported as the maximum bin
value, $b_B$, with an effective upper bound of $\theta_{B+1} = +\infty$.  We
can define all thresholds compactly as \cite{Liddell2018}
\begin{equation}
\theta_j = \begin{cases}
  -\infty         & \text{ if } j=0 \\
  \infty          & \text{ if } j=B \\
  b_1 + \frac{j + 1}{2} & \text{ otherwise.}
\end{cases}
\label{eq:thresholds}
\end{equation}
\noindent

We model the measurement of ordinal opinions by calculating the probability
density function over bins, written
\begin{equation}
  f[b_i] = \int_{\theta_{i - 1}}^{\theta_i} f(\omega) d\omega~,
  \label{eq:measurement-math}
\end{equation}
\noindent
switching to square brackets for the argument to $f$ now to emphasize that 
bins are discrete. With this ordinal measurement model we can caluclate the 
mean opinion of the simulated measurements like so
\begin{equation}
  \bar{o}_t = \frac{1}{B} \sum_{b=b_1}^{b_B} b \cdot f_t[b].
\end{equation}
\noindent

\begin{center}
\vspace{1.5em}
[ FIGURE \ref{fig:distros} ABOUT HERE ]
\vspace{1.5em}
\end{center}

Group polarization is the difference between the post-deliberation mean
observed opinion and the pre-deliberation one, 
which we can write as 
\begin{equation}
  g = \bar{o}_\mathrm{post} - \bar{o}_\mathrm{pre}.
  \label{eq:group-polarization}
\end{equation}
\noindent

\begin{center}
\vspace{1.5em}
[ TABLE \ref{tab:simulated-observations} ABOUT HERE ]
\vspace{1.5em}
\end{center}

\subsection{Root-finding Algorithm to Find Null-Polarization Models with Artifacts}

\subsection{Corpus Analysis to Determine Undecidable Replications}

To analyze published replications we need to expand our notation to index
individual \emph{experiments} with replications, and from which 
different journal articles, track and analyze it,

\subsection{Evaluating Grouop Polarization Experimental Designs}

>>>>

\section{Method}

A group polarization replication is proven undecidable, and therefore
not truly a replication, if we can find a model of \emph{mere agreement}
that appears to be polarized under that replication's experimental
design. If we cannot find such a model, we cannot determine a
replication to be undecidable. Mere agreement in this model is a dynamic
where agreement increases within the group, but extremism does not---in
other words, A \emph{null-polarization distributions} is a set of two
Gaussian distributions: one representing the pre-discussion opinions of
the group, and the other representing the post-discussion opinions. Our
\emph{null-polarization distributions} represent the process of the
group finding \emph{mere agreement}, which is when the average group
opinion is constant over time, but opinion variance decreases to
represent agreement (or, equivalently, \emph{consensus}~(DeGroot 1974)).

\subsection{Latent Opinion Model}\label{latent-opinion-model}

Throughout this paper we model a participant \(i\)'s latent opinion at
time \(t\) as a normally-distributed random variable,
\begin{equation}\phantomsection\label{eq-random-variable-def}{ 
\omega_{i,t} \sim
\mathcal{N}(\mu_t, \sigma_t),
}\end{equation} with mean \(\mu_t\) and standard deviation \(\sigma_t\).
Group polarization in terms of latent opinions is then defined as any
pair of pre- and post-discussion opinion distributions where the mean
increases in magnitude over time---formally,
\(| \mu_\mathrm{post}| > | \mu_\mathrm{pre}|\). We need to take
magnitudes since radicalization may be towards increasingly negative or
positive opinion values.

For group polarization experiments, only two times are relevant:
\emph{pre-discussion}, \(t=\mathrm{pre}\), and \emph{post-discussion},
\(t=\mathrm{post}\). Since participant groups are assumed to be
likeminded, we could also assume they come to agree more over time as
well as become more extreme in our group polarization model, or in terms
of the latent opinion variance,
\(\sigma_\mathrm{pre}> \sigma_\mathrm{post}\)---but only one study in
our ten actually measured opinion variance to check this (Schkade,
Sunstein, and Hastie 2010).

A \emph{null-polarization model} is a model that is consistent with the
\emph{negation}, or \emph{null}, of the \emph{group polarization
hypothesis}. The group polarization hypothesis is that group discussion
causes the average group opinion to increase, assuming the group is
aligned on a particular issue, though opinion strength varies. Therefore
a null model is one where the latent mean stays constant, i.e.,
\(\mu = \mu_\mathrm{pre} = \mu_\mathrm{post}\).

\emph{Mere agreement} is the type of null-polarization model we use
here. Mere agreement is a null-model because the mean latent opinion is
set to be constant before and after discussion. We represent
\emph{agreement}---or equivalently, \emph{consensus}---as a decrease in
the standard deviation that defines the pre- and post-discussion latent
opinion distributions. The mere agreement model is the cornerstone of
our analysis, since it is empirically plausible and generates
\emph{spurious} observations of group polarization under mere agreement
models, due to measurement artifacts induced by clipping all latent
opinion values outside the measurement range---we call this
\emph{spurious polarization} for short. We simulate opinion measurements
in two different ways, but both use the mere agreement model of latent
opinions to test whether and by how much the measurement procedure
distorts null-polarization to look like group polarization.

{[} TABLE
{[}tab:latent-variables{]}(\#tab:latent-variables)\{reference-type=``ref''
reference=``tab:latent-variables''\} ABOUT HERE {]}

\subsection{Two Models of Group Polarization
Measurement}\label{two-models-of-group-polarization-measurement}

We simulate measurement in two different ways---i.e., with two different
models---for different parts of our analysis. The first is the
\emph{expectation model}: an efficient way to \emph{simulate measurement
of the observed average} opinion of a group, using the
\emph{mathematical expectation value} of the distribution of measured
opinions as a proxy. In the second model, the \emph{stochastic model},
we simulate individual participant responses by first drawing an opinion
for each participant from the latent opinion distribution
(Equation~\ref{eq-random-variable-def}). The opinions are binned to the
appropriate ordinal value.

We use the expectation model for its efficiency in the first step of our
analysis: to find a \emph{null-polarization model} that also fits the
data for each experimental condition studied---if we find one that means
we must retract this replication since this proves that the group
polarization is undecidable with the data we have. Once we find any such
undecidable results, we use the null-polarization model to help us
further evaluate the experimental design itself.

In the \emph{stochastic model}, we simulate group polarization data for
the \emph{Evaluation} step of the analysis. We simulate experimental
data, with the same number of participants as the original experiment,
using the null-polarization model found in the \emph{Identification}
step, if one was found.

We \emph{evaluate} the experimental design, \(e\), in terms of the false
positive rate, also known as the Type I error rate or family-wise error
rate, denoted \(\alpha_e\). We close our analysis by demonstrating how
our method can be inverted to \emph{evaluate} experimental designs in
the pre-registration stage to avoid preventable replication failures
like we identify here.

The measurement models are introduced in detail along with the
corresponding stage of the analysis that uses them. First, the
\emph{Identification} stage that uses the \emph{expectation} measurement
model; second, the \emph{Evaluation} stage that uses the
\emph{stochastic} measurement model.

\subsection{\texorpdfstring{Part 1: \emph{Identify} Invalid
Replications}{Part 1: Identify Invalid Replications}}\label{part-1-identify-invalid-replications}

In the first stage of our analysis, we try to \emph{identify} a null
polarization models, \(\nu\), that invalidate published detections of
group polarization if the measurement procedure induces \emph{spurious
group polarization} for that null polarization model. Null polarization
can take many forms---all that is required is that the latent mean
opinion stays constant before and after discussion. Here we use a
\emph{mere agreement} null polarization model, where we simulate
\emph{consensus}, the tendency for sufficiently similar opinions to
become more similar---but we leave out polarization. In mere agreement,
moderate and extreme opinions approach change by equal magnitude towards
each other. We therefore represent consensus as a decrease in the
standard deviation of the latent opinion distribution from pre- to
post-discussion. We simulate the measurement of the group's average
opinion before and after discussion and measure the spurious group
polarization, if there is any. We use this procedure in a search
algorithm we developed to identify which combination of constant latent
mean and pre- and post-discussion opinion variances generate spurious
group polarization, if there are any.

Our invalidation strategy, then, is to find a mere agreement null
polarization model, \(\nu\), that predicts the exact observations
reported in each experimental replication in our corpus. If one is
found, then we have found a model consistent with the \emph{negation} of
the group polarization hypothesis that explains the data as well as the
group polarization hypothesis itself. In other words, the hypothesis is
not proved or disproved given the data, but \emph{undecidable}: the data
cannot \emph{decide} the hypothesis one way or another because of
potential measurement artifacts. Because the experiment failed to
produce a valid replication, the results are effectively stripped of
evidentiary value and warrant exclusion from the aggregate evidence
base. We apply our analysis to every experimental condition across all
ten journal articles in our corpus, which we label \(e\).

But how do we find a \(\nu\) that generates spurious group polarization?
And how do we even simulate the group polarization experiments and
measurements that make mere agreement in latent opinions look like group
polarization? We simulate experimental measurements of group
polarization by using the mathematical \emph{expectation value} of
latent opinions defined by \(\nu\) as a proxy for the observed average
value of ordinal-valued reported opinions (described in the first
subsection below). We then use this measurement model in a numerical
root-finding algorithm that identifies a \(\nu\) that matches published
observations, if possible. If we find such a \(\nu\), this means we have
a model consistent with the null hypothesis that explains published
observations. In other words, the replication is invalidated since the
group polarization hypothesis is \emph{undecidable} given the
data---i.e., it cannot decide whether the hypothesis or its null is
correct.

So, to review, this subsection introduces how we identify which
experimental data can be explained by a mere agreement null model, where
the ordinal survey measurement procedure induces a clipping artifact
that hides changes in extreme opinions that fall outside the measurement
scale.

\subsubsection{The Expectation Model of Spurious Group
Polarization}\label{the-expectation-model-of-spurious-group-polarization}

We use the \emph{expectation model} of opinion measurement to
\emph{Identify} which, if any, mere agreement null-polarization models
generate spurious group polarization in a particular experiment, \(e\).
The trick to this model, and the source of its name, is that we use the
theoretical \emph{expected value} of opinion measurements as a proxy for
the \emph{observed average} of participant opinions under an
experimental design, \(e\).

We simulate the measurement of the mean observed opinion at time \(t\)
by first calculating integrating the normal probability density
function, \(f(\omega)\), over measurement bin boundaries, over each
bin's corresponding range of opinion values. For example, consider an
ordinal scale that runs from -3 (strongly disagree) to +3 (strongly
agree) (as used by Moscovici and Zavalloni (1969), for example). the
frequency with which the ``2'' bin is chosen is just the total area
underneath the probability density curve between 1.5 and 2.5.

All bins have a width of 1 in latent opinion space, except for the two
most extreme bins at opposite ends of the sentiment spectrum---this
works because bin values are always separated by 1. All continuous
latent opinion values \(\omega < b_1 + \frac{1}{2} = \theta_1\) get
mapped to the ordinal measurement value \(b_1\)---so, \(b_1\)'s lower
threshold is \(\theta_0 = -\infty\). Similarly, all opinions
\(\omega > b_B - 0.5 = \theta_B\) are reported as the maximum bin value,
\(b_B\), with an effective upper bound of \(\theta_{B+1} = +\infty\). We
can define all thresholds compactly as (Liddell and Kruschke 2018)
\begin{equation}\phantomsection\label{eq-bin-thresholds}{
\theta_j = \begin{cases}
  -\infty         & \text{ if } j=0 \\
  \infty          & \text{ if } j=B \\
  b_j + \frac{1}{2} & \text{ otherwise.}
\end{cases}
}\end{equation}

The probability density that a participant reports the opinion in bin
index \(i\) at time \(t\) is \[
f_{it}(\mu, \sigma_t) = \int_{\theta_{i - 1}}^{\theta_{i}} f(\omega; \mu, \sigma_t)~d\omega,
\] which we calculated in R to evaluate the integral at the two
threshold values \(\theta_{i-1}\) and \(\theta_i\). Note that the normal
probability density function we used is parameterized by mean \(\mu\)
and standard deviation \(\sigma_t\). \[
f(\omega; \mu, \sigma_t) = 
    \frac{1}{\sigma_t \sqrt{2 \pi}} e^{-\frac{(\omega - \mu)^2}{2 \sigma_t^2}}.
\]

To simulate the measurement of the \emph{average} observed opinion at
time \(t\), \(\bar{o}_t\), as the expected value of the probability
density function of observed ordinal opinions. In other words, we
simulate \(\bar{o}_t\) by calculating the weighted sum of products
between the probability of observing each bin value and the bin value
itself: \begin{equation}\phantomsection\label{eq-sim-ave-observed}{
\bar{o}_t(\nu, e) = \sum_{i=1}^B b_i f_{it}(\nu, e).
}\end{equation}

Group polarization is the difference between the post-deliberation mean
observed opinion and the pre-deliberation one. We will calculate the
magnitude of spurious group polarization under a mere agreement null
model represented by its collected distribution parameters,
\(\nu = (\mu, \sigma_\mathrm{pre}, \sigma_\mathrm{post})\), under
experimental design, \(e\):
\begin{equation}\phantomsection\label{eq-spurious-gp}{
g(\nu, e) = |\bar{o}_\mathrm{post}(\nu, e)| - |\bar{o}_\mathrm{pre}(\nu, e)|
}\end{equation}

In the \emph{Inspect} stage of our analysis we will determine which
group polarization replications are invalidated due to equifinality
between the \emph{mere agreement null-polarization hypothesis} on the
one hand, and the \emph{group polarization hypothesis} on the other.
This requires more than just showing there is some \(\nu\) that causes
spurious polarization for some experiment where \(g(\nu, e) > 0\). We
have to find a \(\nu\) that generates the exact \(g(\nu, e)\) reported
in the original journal article \emph{and} check that simulated
\(\bar{o}_\mathrm{pre}\) and \(\bar{o}_\mathrm{pre}\) match published
values.

{[} TABLE
\hyperref[tab:experiment-parameters]{{[}tab:experiment-parameters{]}}
ABOUT HERE {]}

{[} TABLE
\hyperref[tab:simulated-observations]{{[}tab:simulated-observations{]}}
ABOUT HERE {]}

\subsubsection{Example: Invalidating the ``Americans'' condition in
Moscovici and Zavalloni
(1969)}\label{example-invalidating-the-americans-condition-in-moscovici-and-zavalloni-1969}

One of the replications we invalidate here is from Moscovici and
Zavalloni (1969), the ``Americans'' condition as they called it---we
coded it \texttt{Moscovici1969\_Americans}. They report that participant
opinions polarized after discussing the survey topic, ``American
economic aid is always used for political pressure.'' They report survey
responses increased in extremity from an average of -0.9 before
discussion to -1.6 after, on a seven-bin scale from -3 indicating
``strongly disagree'' to +3 indicating ``strongly agree''.

Moscovici and Zavalloni say their results confirm the basic hypothesis,
that \emph{group discussion causes polarization}. But we found a mere
agreement null-polarization model that generates spurious group
polarization using our model---specifically,
\(\nu = (\mu = -1.2, \sigma_\mathrm{pre}= 4.3, and \sigma_\mathrm{post}= 1.9)\).
This means that the hypothesis is actually undecidable given the data,
invalidating Moscovici and Zavalloni's replication. When a hypothesis
and its null explain data equally well, the data tell us nothing.

Now let's use the expectation model to prove that this \(\nu\) generates
spurious group polarization in this experiment, i.e.,
\(g(\nu, e= \mathrm{Moscovici1969\_Americans}) > 0\). In the process we
demonstrate how the expectation model works to simulate the measurement
of spurious group polarization (Equation~\ref{eq-spurious-gp}).

First, we calculate the probability density vector using the
\href{https://github.com/mt-digital/gp-statmod/blob/main/GroupPolarizationStatmod/model.R\#L44}{\texttt{makeProbVec}}
function in the file
\href{https://github.com/mt-digital/gp-statmod/blob/main/GroupPolarizationStatmod/model.R}{\texttt{GroupPolarizationStatmod/model.R}}
in the \href{https://github.com/mt-digital/gp-statmod}{project code
repository on GitHub} for both experiment times: pre- and
post-discussion: \[
f_\mathrm{pre} = 
\begin{pmatrix}
  0.38 & 0.09 & 0.09 & 0.09 & 0.08 & 0.07 & 0.19
\end{pmatrix}
\] and \[
f_\mathrm{post} = \
\begin{pmatrix} 
0.25 & 0.19 & 0.21 & 0.17 & 0.11 & 0.05 & 0.03 
\end{pmatrix}.
\]

When we plug in the bin values and copmuted \(f_\mathrm{pre}\) values
into the simulated average pre-discussion opinion definition
(Equation~\ref{eq-sim-ave-observed}), we get \[
\bar{o}_\mathrm{pre} = -3 \cdot 0.38~~+-2\cdot 0.09~~+\ldots+3\cdot0.19 = -0.6
\] and \[
\bar{o}_\mathrm{post} = -3 \cdot 0.25~~+-2\cdot 0.19~~+\ldots+3\cdot0.03 = -1.04.
\] So, we find that for the \texttt{Moscovici1969\_Americans}, our
null-polarization model appears polarized, with the exact same average
pre- and post-discussion opinions reported by Moscovici and Zavalloni
(1969), and the same magnitude of radicalization: \[
g(\nu, e) = \bar{o}_\mathrm{post} - \bar{o}_\mathrm{pre} = -1.04 - (-0.6) = - 0.4.
\]

FIGURE \hyperref[fig:distros]{{[}fig:distros{]}} ABOUT HERE

Since the simulated average opinions match the observed ones---note that
as a consequence, the spurious polarization magnitude also matches the
observed one. In other words, we \emph{identified} a null model that
explains the data. If both a hypothesis and its null explain data
equally well, then the hypothesis is \emph{undecidable} under the
available data. This invalidates this experiment's claim to be a
replication of support for the group polarization hypothesis. We
replicate this analysis for all 57 experimental conditions that purport
to support the group polarization hypothesis. To facilitate this
analysis, we created a web app to manage the data entry, the process of
identifying invalid replications, and managing the outcome data of which
replications were invalidated, and what null model parameters generated
the spurious polarization.

\subsubsection{System for identifying invalid
replications}\label{system-for-identifying-invalid-replications}

The next step is to scale up the analysis we did manually in the example
above, including develop a systematic way to find, and not guess, the
set of null model parameters that

\begin{figure}

\centering{

\includegraphics[width=10.53125in,height=\textheight,keepaspectratio]{Figures/Model/SoftwareMap_Mashup.png}

}

\caption{\label{fig-computational-system}\textbf{Computational system to
interactively simulate opinion change, identify invalid replications,
and evaluate experimental designs.}}

\end{figure}%

\subsection{\texorpdfstring{Part 2: \emph{Evaluate} Experimental
Designs}{Part 2: Evaluate Experimental Designs}}\label{part-2-evaluate-experimental-designs}

In the previous step we identified which, if any, null polarization
models generate spurious group polarization with the expectation model
of opinion measurement. This is enough to invalidate a replication, but
it leaves unanswered questions. First, we would like to know how what
the chance was that a published detection of group polarization was
actually a false positive, assuming that participant opinions were
distributed according to the null model identified in the first part of
the analysis. This is equivalent to the false positive rate of the
experimental design, which we denote \(\alpha_e\) for experiment \(e\).
Second, we would like to estimate the false discovery rate for our
corpus, which is a good proxy for the field as a whole, since we
selected top quality, highly influential work that has been cited
thousands of times. The false discovery rate in this case tells us the
probability that a published detection of group polarization is actually
a false detection. (Equation~\ref{eq-fdr}).

The core of our evaluation is the quantification of the probability that
a spurious result---an effect generated by the null model---is falsely
interpreted as genuine group polarization. We calculate this probability
across three distinct units of analysis, building from the specific to
the general:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  The Type I Error Rate (\(\alpha_e\)): The rate of falsely detecting
  group polarization in a single experimental condition (\(e\)). This
  directly quantifies the risk posed by the mere agreement model
  identified for that specific experiment.
\item
  The study-level Family-Wise Error Rate (\(\alpha_s\)): The mean Type I
  Error Rate aggregated across all conditions within a single study
  (\(s\)). This estimates the false claim rate across the set of tests
  performed in that study.
\item
  The Family-Wise Error Rate (\(\alpha_{\text{all}}\)): The mean Type I
  Error Rate aggregated across all 57 experimental conditions in the
  corpus (\(E\)). This estimates the expected false claim rate for a
  researcher adopting the general experimental design found throughout
  the literature.
\end{enumerate}

These calculated Type I Error Rates (\(\alpha\)) are then combined with
fixed external parameters to estimate the False Discovery Rate.
Following convention, we adopt a fixed Statistical Power of \(W=0.80\)
(and test sensitivity with alternative values) and a Base Rate of
\(b=0.10\). The relationship between the units of analysis and the
estimated error rates is summarized in Table \textbf{?@tbl-error-rates}.

Statistical Error Rates and Units of Analysis

Unit of Analysis

Type

Notation

Definition

Experiment, \(e\)

Type I

\(\alpha_e\)

The rate of falsely rejecting the null hypothesis (\(d=0\)) for that
single test (used in aggregation).

All experiments in a study, \(s\)

Family-Wise

\(\alpha_s\)

The mean Type I Error Rate across all conditions in that specific study.

All Experiments

Family-Wise

\(\alpha_{\text{E}}\)

The mean Type I Error Rate across all conditions identified as plausibly
spurious.

\subsubsection{Generative Model of Spurious Group
Polarization}\label{generative-model-of-spurious-group-polarization}

To evaluate an experimental design's ability to properly identify null
effects, we need to simulate participant responses assuming the mere
agreement null model that invalidated the replication (\(\nu_e\)). We
use the stochastic model for this simulation stage, as it correctly
models the noise inherent in individual participant measurement and
response. The ordinal, measured opinions are generated by executing the
following steps for \(N_e\) simulated participants across \(N_T\)
independent trials:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Latent Opinion Sampling:} For each participant \(i\) and time
  \(t \in \{\text{pre, post}\}\), we draw a continuous latent opinion
  \(\omega_{i,t}\) from the corresponding normal distribution defined by
  the null model \(\nu_e\):
  \[\omega_{i,t} \sim \mathcal{N}(\mu, \sigma_t)\] where \(\mu\) is the
  constant latent mean and \(\sigma_t\) is the time-specific standard
  deviation derived from the mere agreement null model \(\nu_e\).
\item
  \textbf{Measurement Simulation (Ordinal Binning):} Each continuous
  latent opinion \(\omega_{i,t}\) is mapped onto the experiment's
  ordinal scale (e.g., \(-3\) to \(+3\)). This mapping is determined by
  the fixed measurement thresholds, \(\theta_j\), established in the
  \emph{Expectation Model} subsection
  (Equation~\ref{eq-bin-thresholds}). This process simulates the
  clipping artifact that generates spurious polarization in the
  observable data.
\end{enumerate}

This results in a simulated dataset of \(2N_e\) rows (two time steps per
participant) for a single trial, consisting of ordinal opinion values
that are then passed to the statistical inference pipeline.

\subsubsection{Statistical Inference
Pipeline}\label{statistical-inference-pipeline}

To complete our our simulated experimental design we need to be able to
statistically infer the effect sizes that measure the strength of group
polarization in an experiment. To keep the presentation simple and
transparent, we infer effect sizes by inferring the parameters \(\mu_t\)
and \(\sigma_t\) of the hypothesized normal distribution that generates
the data---\emph{literally} generates the data in our simulations, but
is just a model of the distribution of group opinions in real-world
experiments.

\subsubsection{False positive rate}\label{false-positive-rate}

To calculate the false detection rate we begin by simulating pre- and
post-deliberation opinions by drawing \(N_e\) samples from normal latent
opinion distributions that generated spurious group polarization, i.e.,
one sample at each experiment stage for each simulated participant. The
latent opinion distributions are specified with the mere agreement null
model labelled \(\nu_e\).

The family-wise error rate is the fraction of trials with
\(d_{ei} > d^*\), written
\begin{equation}\phantomsection\label{eq-study-aggregate}{
  \alpha_e = \frac{1}{N_T} \sum_i^{N_T} \Theta(d_{ei}; d^*),
}\end{equation} where \[
  \Theta(d_{ei}; d^*) = 
    \begin{cases}
      1 & \text{ if } d_{ei} > d^* \\
      0 & \text{ otherwise.}
    \end{cases}
\] \(d^*\) is the \emph{significance threshold}, which must be specified
by the experimenter. There are three common significance levels which we
adopt here: \(d^* = 0.2\) for ``low'', \(d^*=0.5\) for ``medium'', and
\(d^* = 0.8\) for ``high''.

To calculate the family-wise error rate across all experimental
conditions of a study, \(s\), \[
\alpha_s = \frac{1}{|s|} \sum_{e \in s} \alpha_e,
\] where \(e \in s\) indicates sum over all experimental conditions
\(e\) presented in the article associated with study \(s\). The mean
across all results identified as plausibly spurious is \[
 \alpha_\mathrm{all} = \frac{1}{|E|} \sum_{e \in E} \alpha_e.
\]

In the final part of our \emph{Evaluation} analysis we adapt this
analysis to re-align the question of statistical rigor: we can be
specific now, and instead of wondering, ``what significance level is
truly rigorous'', we can use our techniques to make the choice of
\(d^*\) to conform to a specific design requirement. Specifically we
will calculate what \(d^*\) must be in order to limit
\(\alpha \leq 0.05\) as an example.

\subsubsection{False discovery rate}\label{false-discovery-rate}

The false discovery rate is the false positive rate, scaled by the rate
of true negative effects and normalized by the frequency of true
positive effects\textasciitilde{} (Equation~\ref{eq-fdr}) and assuming
empirically-motivated values for base rate, \(b\), and statistical
power, \(W\), as well as more pessimistic and optimistic settings for
base rate and power. Using this notation, the false detection rate is

\begin{equation}\phantomsection\label{eq-fdr}{
\mathrm{FDR}= \frac{(1 - b)\alpha}{(1 - b)\alpha + bW}.
}\end{equation}

\subsection{\texorpdfstring{Part 3: \emph{Specify} Significance Threshold
for Experimental
Design}{Part 3: Specify Significance Threshold for Experimental Design}}\label{part-3-specify-significance-threshold-for-experimental-design}

In the final stage of our analysis, we pivot from evaluating the outcome
of published designs to \emph{specifying the statistical requirements
for future designs}. The primary goal of this specification is to invert
the standard question of statistical rigor: instead of asking if an
effect is `significant' given a fixed effect-size threshold (e.g.,
\(d^{*} = 0.5\)), we ask what significance threshold is necessary to
limit the false positive rate to a desired maximum, such as \(5\%\) (or
\(\alpha = 0.05\)).

We leverage the evaluation framework developed in Part 2, based on the
mere agreement null model \(\nu_e\) identified for each experimental
condition \(e\). Our objective is to calculate the function
\(\alpha_e(d^*)\) across a sufficient range of \(d^*\) values. This
allows us to determine the lowest \(d^*\) that satisfies the design
requirement \(\alpha_e(d^*) \leq 0.05\).

The function \(\alpha_e(d^*)\) is the simulated false positive rate as a
function of the significance threshold, \(d^*\), defined as: \[
\alpha_e(d^*) = \frac{1}{N_T} \sum_i^{N_T} \Theta(d_{ei}; d^*),
\] where \(N_T\) is the number of null trials, \(d_{ei}\) is the
calculated effect size for trial \(i\), and \(\Theta\) is the indicator
function that flags trials where the effect size \(d_{ei}\) exceeds the
significance threshold \(d^*\). In the interest of giving the most
generous best-case scenario possible, we only count positive effect
sizes because only these indicate increased extremism over time.
Negative effect sizes indicate group opinions became more
moderate---such observations do not support the group polarization
hypothesis that opinions become more extreme following discussion.

Because our system is automated, and the code publicly available, it can
be adapted and used to design future group polarization experiments with
empirically-grounded statistical rigor. Experimenters can specify a
desired maximum false positive rate, such as \(\alpha = 0.05\), and our
system will calculate the minimum effect size threshold \(d^*\) that
must be exceeded to achieve that level of rigor under the mere agreement
null model identified for their experimental design. This prescriptive
step transforms the simulation from a post-hoc critique into a
forward-looking design tool, giving experimenters an empirically-derived
minimum effect size they must exceed to maintain statistical confidence
in their results.

\subsection{Analysis Outline and
Summary}\label{analysis-outline-and-summary}
\addcontentsline{toc}{section}{Analysis Outline and Summary}

%\phantomsection\label{refs}
%\begin{CSLReferences}{1}{0}
%\bibitem[\citeproctext]{ref-DeGroot1974}
%DeGroot, Morris H. 1974. {``{Reaching a Consensus}.''} \emph{Journal of
%the American Statistical Association} 69 (345): 118--21.
%
%\bibitem[\citeproctext]{ref-Liddell2018}
%Liddell, Torrin M., and John K. Kruschke. 2018. {``{Analyzing ordinal
%data with metric models: What could possibly go wrong?}''} \emph{Journal
%of Experimental Social Psychology} 79: 328--48.
%\url{https://doi.org/10.1016/j.jesp.2018.08.009}.
%
%\bibitem[\citeproctext]{ref-Moscovici1969}
%Moscovici, Serge, and Marisa Zavalloni. 1969. {``{The group as a
%polarizer of attitudes}.''} \emph{Journal of Personality and Social
%Psychology} 12 (2): 125--35.
%
%\bibitem[\citeproctext]{ref-Schkade2010}
%Schkade, David, Cass R. Sunstein, and Reid Hastie. 2010. {``{When
%deliberation produces extremism}.''} \emph{Critical Review} 22 (2):
%227--52. \url{https://doi.org/10.1080/08913811.2010.508634}.
%
%\end{CSLReferences}
%
%
%
%
%\end{document}
